maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-07-27 docs/paper/paper.tex
add some DOIs
more DOIs
fix umlaut in paper?

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={N/A},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[4%
%
]{Aaron Sandoval%
  \,\orcidlink{0009-0002-8380-6140}\,%
}
\author[2%
%
]{Alex F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[3%
%
]{Tilman Räuker%
  \,\orcidlink{0009-0009-6321-4413}\,%
}
\author[1%
%
]{Brandon Knutson%
  \,\orcidlink{0009-0004-8413-0239}\,%
}
\author[1%
%
]{Cecilia Diniz Behn%
  \,\orcidlink{0000-0002-8078-5105}\,%
}
\author[1%
%
]{Samy Wu Fung%
  \,\orcidlink{0000-0002-2926-4582}\,%
}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
1.0: A complete set of tools to prepare, present, and analyze
navigational and spatial cognitive neuroscience experiments.
\emph{Behavior Research Methods}, \emph{40}, 353--359.

\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.

\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
procedural generation to benchmark reinforcement learning. \emph{arXiv
Preprint arXiv:1912.01588}.

\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
models' spatial intelligence via GRPO. \emph{arXiv Preprint
arXiv:2502.14669}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
(2022). Jfb: Jacobian-free backpropagation for implicit networks.
\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
\emph{36}, 6648--6656.

@@

\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
{Games CoG}}, 1--4.

\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
Ivanitskiy, M. (n.d.). \emph{ZANJ}.
\url{https://github.com/mivanit/ZANJ}

\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
\emph{Maze dataset}.
\url{https://github.com/understanding-search/maze-dataset}

\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
\emph{Maze transformer interpretability}.
\url{https://github.com/understanding-search/maze-transformer}

\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
C. D., \& others. (2023). A configurable library for generating and
manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.

\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
D., \& others. (2023). Structured world representations in maze-solving
transformers. \emph{arXiv Preprint arXiv:2312.02566}.

\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
Ibrahim, M. (2024). The factorization curse: Which tokens you predict
underlie the reversal curse and more. \emph{Advances in Neural
Information Processing Systems}, \emph{37}, 112329--112355.

\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
extrapolation for mazes with recurrent and implicit networks.
\emph{arXiv Preprint arXiv:2410.03020}.

\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
Performance insights and comparative analysis. \emph{arXiv Preprint
arXiv:2308.11224}.

\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
Nag, A. (2020). MDL suite: A language, generator and compiler for
describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
1815.

@@

\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
Transformers can navigate mazes with multi-step prediction. \emph{arXiv
Preprint arXiv:2412.05117}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
Hidden computation in transformer language models. \emph{arXiv Preprint
arXiv:2404.15758}.

\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
transparent ai: A survey on interpreting the inner structures of deep
neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
Machine Learning (SaTML)}, 464--483.

@@

\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
Generalizing from easy to hard problems with recurrent networks.
\emph{Advances in Neural Information Processing Systems}, \emph{34},
6695--6706.

\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
\emph{arXiv Preprint arXiv:2308.02522}.

\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
world models in maze-solving tasks. \emph{arXiv Preprint
arXiv:2412.11867}.

\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
self-supervised neural-symbolic learning framework for robot autonomy.
\emph{arXiv Preprint arXiv:2406.16087}.

\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
arXiv:2502.01989}.

>>>>

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Tilman R"auker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={N/A},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Tilman R"auker, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[4%
%
]{Aaron Sandoval%
  \,\orcidlink{0009-0002-8380-6140}\,%
}
\author[2%
%
]{Alex F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[3%
%
]{Tilman R"auker%
  \,\orcidlink{0009-0009-6321-4413}\,%
}
\author[1%
%
]{Brandon Knutson%
  \,\orcidlink{0009-0004-8413-0239}\,%
}
\author[1%
%
]{Cecilia Diniz Behn%
  \,\orcidlink{0000-0002-8078-5105}\,%
}
\author[1%
%
]{Samy Wu Fung%
  \,\orcidlink{0000-0002-2926-4582}\,%
}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ayaz2008maze}{}}%
Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
1.0: A complete set of tools to prepare, present, and analyze
navigational and spatial cognitive neuroscience experiments.
\emph{Behavior Research Methods}, \emph{40}, 353--359.
\url{https://doi.org/10.3758/brm.40.1.353}

\leavevmode\vadjust pre{\hypertarget{ref-chen2024iaimperative}{}}%
Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
\url{https://doi.org/10.48550/arXiv.2403.15870}

\leavevmode\vadjust pre{\hypertarget{ref-cobbe2019procgen}{}}%
Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
procedural generation to benchmark reinforcement learning. \emph{arXiv
Preprint arXiv:1912.01588}.
\url{https://doi.org/10.48550/arXiv.1912.01588}

\leavevmode\vadjust pre{\hypertarget{ref-dao2025alphamaze}{}}%
Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
models' spatial intelligence via GRPO. \emph{arXiv Preprint
arXiv:2502.14669}. \url{https://doi.org/10.48550/arXiv.2502.14669}

@@

\leavevmode\vadjust pre{\hypertarget{ref-fung2022jfb}{}}%
Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
(2022). Jfb: Jacobian-free backpropagation for implicit networks.
\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
\emph{36}, 6648--6656. \url{https://doi.org/10.1609/aaai.v36i6.20619}

@@

\leavevmode\vadjust pre{\hypertarget{ref-harriesMazeExplorerCustomisable3D2019}{}}%
Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
{Games CoG}}, 1--4. \url{https://doi.org/10.1109/cig.2019.8848048}

\leavevmode\vadjust pre{\hypertarget{ref-zanj}{}}%
Ivanitskiy, M. (n.d.). \emph{ZANJ}.
\url{https://doi.org/10.5281/zenodo.15540393}

\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-github}{}}%
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
\emph{Maze dataset}. \url{https://doi.org/10.48550/arXiv.2309.10498}

\leavevmode\vadjust pre{\hypertarget{ref-maze-transformer-github}{}}%
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
\emph{Maze transformer interpretability}.
\url{https://doi.org/10.48550/arXiv.2312.02566}

\leavevmode\vadjust pre{\hypertarget{ref-maze-dataset-arxiv-2023}{}}%
Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
C. D., \& others. (2023). A configurable library for generating and
manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
\url{https://doi.org/10.48550/arXiv.2309.10498}

\leavevmode\vadjust pre{\hypertarget{ref-ivanitskiy2023structuredworldreps}{}}%
Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
D., \& others. (2023). Structured world representations in maze-solving
transformers. \emph{arXiv Preprint arXiv:2312.02566}.
\url{https://doi.org/10.48550/arXiv.2312.02566}

\leavevmode\vadjust pre{\hypertarget{ref-MLMU-kitouni2024factorization}{}}%
Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
Ibrahim, M. (2024). The factorization curse: Which tokens you predict
underlie the reversal curse and more. \emph{Advances in Neural
Information Processing Systems}, \emph{37}, 112329--112355.
\url{https://doi.org/10.48550/arXiv.2406.05183}

\leavevmode\vadjust pre{\hypertarget{ref-knutson2024logicalextrapolation}{}}%
Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
extrapolation for mazes with recurrent and implicit networks.
\emph{arXiv Preprint arXiv:2410.03020}.
\url{https://doi.org/10.48550/arXiv.2410.03020}

\leavevmode\vadjust pre{\hypertarget{ref-eval-LLM-graphs}{}}%
Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
Performance insights and comparative analysis. \emph{arXiv Preprint
arXiv:2308.11224}. \url{https://doi.org/10.48550/arXiv.2308.11224}

\leavevmode\vadjust pre{\hypertarget{ref-mdl-suite}{}}%
Nag, A. (2020). MDL suite: A language, generator and compiler for
describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
1815. \url{https://doi.org/10.21105/joss.01815}

@@

\leavevmode\vadjust pre{\hypertarget{ref-nolte2024multistep}{}}%
Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
Transformers can navigate mazes with multi-step prediction. \emph{arXiv
Preprint arXiv:2412.05117}.
\url{https://doi.org/10.48550/arXiv.2412.05117}

@@

\leavevmode\vadjust pre{\hypertarget{ref-pfau2024dotbydot}{}}%
Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
Hidden computation in transformer language models. \emph{arXiv Preprint
arXiv:2404.15758}. \url{https://doi.org/10.48550/arXiv.2404.15758}

\leavevmode\vadjust pre{\hypertarget{ref-interpretability-survery}{}}%
Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
transparent ai: A survey on interpreting the inner structures of deep
neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
Machine Learning (SaTML)}, 464--483.
\url{https://doi.org/10.1109/satml54575.2023.00039}

@@

\leavevmode\vadjust pre{\hypertarget{ref-deepthinking}{}}%
Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
Generalizing from easy to hard problems with recurrent networks.
\emph{Advances in Neural Information Processing Systems}, \emph{34},
6695--6706. \url{https://doi.org/10.48550/arXiv.2106.04537}

\leavevmode\vadjust pre{\hypertarget{ref-eval-gpt-visual}{}}%
Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
\emph{arXiv Preprint arXiv:2308.02522}.
\url{https://doi.org/10.48550/arXiv.2308.02522}

\leavevmode\vadjust pre{\hypertarget{ref-spies2024causalworldmodels}{}}%
Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
world models in maze-solving tasks. \emph{arXiv Preprint
arXiv:2412.11867}. \url{https://doi.org/10.48550/arXiv.2412.11867}

\leavevmode\vadjust pre{\hypertarget{ref-wang2024imperative}{}}%
Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
self-supervised neural-symbolic learning framework for robot autonomy.
\emph{arXiv Preprint arXiv:2406.16087}.
\url{https://doi.org/10.48550/arXiv.2406.16087}

\leavevmode\vadjust pre{\hypertarget{ref-zhang2025tscend}{}}%
Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025). T-SCEND: Test-time
scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
arXiv:2502.01989}. \url{https://doi.org/10.48550/arXiv.2502.01989}
